Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 arch.tex
Edits to dropping policies section.
Add example to architecture section.

% Overview
In order to perform filtering, we need to be able to identify and track flows
of null metadata. Similarly, for partial monitoring 
, we need to be able to track the flow of invalid metadata. In
order to efficiently these flows, we use a dedicated hardware
dataflow engine. A block diagram of the dataflow engine is shown in
Figure~\ref{fig:arch.dataflow}.

@@

\subsection{Example Extensions}

>>>>

% Overview
In order to perform filtering, we need to be able to identify and track flows
of null metadata. Similarly, for partial monitoring,
 we need to be able to track the flow of invalid metadata. In
order to efficiently these flows, we use a dedicated hardware
dataflow engine. A block diagram of the dataflow engine is shown in
Figure~\ref{fig:arch.dataflow}.

@@

\subsection{Example Monitoring Scheme}

% Monitoring operations
\begin{table*}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/monitor}
    \end{small}
    \caption{Monitoring operations for array bounds check.}
    \label{tab:arch.monitor}
    \vspace{-0.2in}
  \end{center}
\end{table*}

% Example dataflow operations
\begin{table*}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/dataflow_operations}
    \end{small}
    \caption{Filtering conditions and operations for clean flags.}
    \label{tab:arch.dataflow_operations}
    \vspace{-0.2in}
  \end{center}
\end{table*}

In this section, we show an example implementation of how the dataflow engine
is configured for an array bounds check monitoring scheme.
Table~\ref{tab:arch.monitor} shows a summary of the monitoring operations for
array bounds check.  When an array pointer is allocated, the metadata
associated with it set to include the base (start) address and bound (end)
address of the corresponding array. On ALU instructions, this base and bound
metadata is copied to the destination registers. On store instructions, the
metadata for the source register is copied to the metadata for the store
address. Similarly, on load instructions, the metadata value for the load
address is copied to the destination register's metadata. In addition, on load
and store instructions, the memory address is checked to be within the base and
bounds metadata. If it is not, then an out-of-bounds error is detected.

Table~\ref{tab:arch.dataflow_operations} summarizes when when monitoring events
can be filtered out due to clean and invalid metadata. In addition, it shows
what metadata needs to be modified on a filtered or dropped monitoring event.
Note that for simplicity the policies for clean and invalid flags are shown
separately. However, the dataflow engine actually operates on pairs of these
flags using a combined policy.

In terms of filtering using clean flags, array allocations are never filtered
since these will always involve non-clean metadata (i.e., setting new base and
bounds). For ALU operations, if both source operands are clean, then the event
can be filtered by setting the destination register as clean. For load and
store, only a single source needs to be checked to be clean in order for the
event to be filtered. Similarly, the destination register or memory location
needs to be marked as clean if the event is filtered.

The invalid flags for array bounds check operate similarly to the clean flags.
When ALU, load, or store events are dropped or filtered, the destination
register or memory location is marked as invalid. In addition, we need to
handle the possibility of dropping an array allocation. In this case, the array
pointer is marked as invalid. In terms of filtering out invalid events, for an
ALU event, if either of the source registers is invalid then the event can be
filtered. For load and store instructions, if the single source operand is
invalid, then the event can be filtered.

\subsection{Applicability to Other Schemes}

Our approach can be generally applied to a wide range of monitoring schemes. This
is possible because the use of clean and invalidation flags is largely agnostic
to the semantics of the specific monitoring scheme. That is, the dataflow
engine must be configured based on the dataflow of the monitoring scheme, but
it does not need detailed information of what the monitoring scheme does.

We do note some limitations to the proposed approach. Previous work
\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
overheads has also filtered out redundant updates where the source metadata
matches the destination metadata. Since we do not read the actual metadata, our
architecture is not able to identify these redundant updates. However, by using
only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
In addition, we note that filtering using clean flags works well when a large
portion of instructions are not relevant to the monitoring scheme (e.g., array
bounds check only cares about array pointers). However, if most instruction
have dirty metadata, then very little filtering will be possible.

In terms of using invalidation to perform partial monitoring, this 
approach works best for monitoring schemes where there are many pieces of
independent metadata. Since marking a piece of metadata as invalid will cause
future metadata with dependences to also be marked as invalid, schemes with
large metadata dependency chains can quickly see large amounts of metadata
become invalid.

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination.  For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a clean and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as clean may indicate that the entire data structure's metadata
is clean.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support more complex monitoring schemes.

### 2014-05-30 tabs/drop.tex
Add example to architecture section.

\multirow{4}{*}{BC}  
& Array allocation & Never & Invalidate array pointer \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\
\hline\hline

\multirow{2}{*}{UMC}  
& Store & Never & Invalidate destination memory location \\ \cline{2-4}
& Load & Invalid source memory location & Do nothing \\ 
\hline\hline

\multirow{5}{*}{DIFT}  
& Software set tag & Never & Invalidate destination memory location \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\ \cline{2-4}
& Indirect jump & Invalid source metadata & Do nothing \\ 
\hline

>>>>

\multirow{4}{*}{BC}  
& Array allocation & Never & Invalidate array pointer \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\
\hline%\hline

% \multirow{2}{*}{UMC}  
% & Store & Never & Invalidate destination memory location \\ \cline{2-4}
% & Load & Invalid source memory location & Do nothing \\ 
% \hline\hline
% 
% \multirow{5}{*}{DIFT}  
% & Software set tag & Never & Invalidate destination memory location \\ \cline{2-4}
% & ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
% & Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
% & Store & Invalid source register & Invalidate destination memory location \\ \cline{2-4}
% & Indirect jump & Invalid source metadata & Do nothing \\ 
% \hline

### 2014-05-30 tabs/monitor.tex
Add example to architecture section.

\begin{tabular}{|l|l|l|}
\hline

{\bf Monitor} & {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline

\multirow{4}{*}{BC}  
 & Array allocation & Set base and bound metadata \\ \cline{2-3}
 & ALU & Copy metadata from register to register \\  \cline{2-3}
 & Load & Copy metadata from memory to register. Check memory access address. \\ \cline{2-3}
 & Store & Copy metadata from register to memory. Check memory access address. \\ 
\hline\hline

\multirow{2}{*}{UMC}  
 & Store & Set metadata tag (data initialized) \\ \cline{2-3}
 & Load & Check if metadata tag is set \\ 
\hline\hline
% & Software clear tag & Clear metadata tag & Non-critical \\ \hline\hline

\multirow{5}{*}{DIFT}  
 & Software set tag & Set tag to a specified value based on software command \\ \cline{2-3}
 & ALU & Output tag is the OR of the input tags \\ \cline{2-3}
 & Load & Set register tag as tag of loaded data \\ \cline{2-3}
 & Store & Set memory tag as register tag \\ \cline{2-3}
% & Software clear tag & Clear tag & Update \\ \cline{2-4}
 & Indirect jump & Check tag \\ 
 \hline

% \multirow{5}{*}{BC}  
%  & MOV/LD/ST          & Propagate colors (Tag[dest] = Tag[src]) & Critical \\ \cline{2-4}
%  & ADD/SUB/NOT        & Update colors (Tag[dest] = f(Tag[src])) & Critical \\ \cline{2-4}
%  & OR/XOR             & Clear colors & Critical \\ \cline{2-4}
%  & LD/ST              & Compare pointer and memory colors & Check \\ \cline{2-4}
%  & Software set color & Set pointer colors & Critical \\ \hline\hline

% Soft Error Check & Instruction & Check instruction computation & Check \\ \hline\hline
% 
% \multirow{2}{*}{Bank of Observers}  
%  & System output & Update state estimation & Critical \\ \cline{2-4}
%  & New state estimate & Check that states of all observers matches & Check \\ \hline

>>>>

\begin{tabular}{|l|l|}
\hline

 {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline

 Array allocation & Set base and bound metadata \\ \hline
 ALU & Copy metadata from register to register \\  \hline
 Load & Copy metadata from memory to register. Check memory access address. \\ \hline
 Store & Copy metadata from register to memory. Check memory access address. \\ 
\hline%\hline

% \multirow{2}{*}{UMC}  
%  & Store & Set metadata tag (data initialized) \\ \cline{2-3}
%  & Load & Check if metadata tag is set \\ 
% \hline\hline
% 
% \multirow{5}{*}{DIFT}  
%  & Software set tag & Set tag to a specified identifier value based on software command \\ \cline{2-3}
%  & ALU & If input tags are different, output tag is a new identifier. Otherwise, output tag is the input tag. \\ \cline{2-3}
%  & Load & Set register tag as tag of loaded data \\ \cline{2-3}
%  & Store & Set memory tag as register tag \\ \cline{2-3}
%  & Indirect jump & Check tag \\ 
%  \hline
